---
title: "Part 2 - Intro to time series | Wrangling and vis"
author: "Peyton Moore"
date: "1/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)
```


### Read in data 
```{r}

toolik <- read_csv(here::here("data", "toolikweather.csv"))%>% 
  mutate(date = mdy(date))


```

### Convert to tsibble 

```{r}
toolik_ts <- toolik %>% 
  as_tsibble(key = NULL, 
             index = date)
```


```{r}
ggplot(data = toolik_ts, 
       aes(x = date, 
           y = mean_airtemp))+
  geom_line()
         
```

## Use `index_by()` function to aggregat time iformation by intervals we specify 

```{r}
toolik_month <- toolik_ts %>% 
  index_by(yr_mo = ~yearmonth(.)) %>% 
  summarize(monthly_mean_temp = mean(mean_airtemp, na.rm = TRUE))

ggplot(data = toolik_month, aes(x = yr_mo, 
                                y  = monthly_mean_temp))+
  geom_line()+
  facet_wrap(~month(yr_mo, label = TRUE))
  
```

Aggregate by week: 

```{r}
toolik_weekly <- toolik_ts %>% 
  index_by(weekly = ~yearweek(.)) %>% 
  summarize(weekly_airtemp = mean(mean_airtemp, na.rm = TRUE), 
            min_airtemp = min(mean_airtemp))


head(toolik_weekly)
```

### Filtering Time series 


Use `filter_index()` to specify ranges of time series to include/exclude 

```{r}
toolik_ts %>% 
  filter_index("2000-06" ~ "2001-10")

```


```{r}
toolik_ts %>% 
  filter_index("2006-04-10"~"2006-05-15")
```

### Seasonplots 

Wraps functions, and adds things to ggplot

```{r}
toolik_ts %>%  
  filter(year(date)> 2014) %>% 
  gg_season(y = mean_airtemp)


toolik_month %>% 
  gg_season(y = monthly_mean_temp)+
  theme_light()



```
### Seasonal subseries plot 

```{r}
toolik_month %>% 
  gg_subseries(monthly_mean_temp)
```

### Find moving average
Using the slider package in R 

```{r}
set.seed(2021)

test <- rnorm(n = 100, mean = 40, sd = 10)


slide(test, ~.x, .before = 2, .after = 2)


```



```{r}
roll_toolik_15 <- toolik_ts %>% 
  mutate(ma_15d = as.numeric(slide(toolik_ts$mean_airtemp, 
                                   mean, 
                                   .before = 7, 
                                   .after = 7)))


ggplot()+
  geom_line(data = toolik_ts, aes(x = date, y = mean_airtemp), 
            size = 0.2, 
            color = "gray")+
  geom_line(data = roll_toolik_15, aes(x = date, 
                                       y = ma_15d))+
  theme_minimal()
```
## Create the ACF

```{r}

toolik_ts %>% 
  ACF(mean_airtemp) %>% 
  autoplot()

toolik_month %>% 
  ACF(monthly_mean_temp) %>% 
  autoplot()

```

### Time series decomposition by STL

```{r}
toolik_decomp <- toolik_month %>% 
  model(STL(monthly_mean_temp ~ season(window = Inf)))

head(toolik_decomp)


components(toolik_decomp) %>%  
  autoplot()
```

